---
title: "ESM_205_Lab_5"
author: "Joanne Pham"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Lab 5 objectives

- Learn to prase dates with 'lubridate'
- Get counts of observations with count()
- Uncount() rows
- One and two-smaple t-test
- Create a heatmap wiht geom_tile()

```{r, include = FALSE}
# Read in data and attach packages
library(tidyverse)
library(here)
library(janitor)

# This data below is not tidy because each observation (ie lobster_count) is not in its own row. After tidy, then can be grouped or counted. 

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"), na = "-99999") %>%   
  janitor::clean_names()

```

Use tidyr::uncount() function to convert our lobster data from frequency fromat to case format

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
```
 
 #### Exploratory data vizualizations
 
 We're only going to consider "site" as our variable of interest
 
```{r}

ggplot(lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              alpha = .5, 
              width = .2)

# histogram
ggplot(lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site)

# regardless of what the last two plots showed, as long as we use the central limit theorom, we can still run parametric tests on the means since we assume that the means follow a normal distribution as long as we have more than 30 samples 
```

#### Convert the 'date' column to class 'Date'

We'll use hte 'lubridate' package to convert to Date format, and then to help us easily parse month and year 

```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```
 
Now let's parse year and month using lubridate::month() and lubridate_year()

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)
  )
```

Now let's find counts of observed lobsters based on different hierarchical groupings:

First, let's count lobsters by year and month:
```{r}
lobster_ym <- lobster_parse_date %>% 
  dplyr::count(obs_year, obs_month)

lobster_ym

lobster_y <- lobster_parse_date %>% 
  dplyr::count(obs_year)

lobster_y

lobster_site <- lobster_parse_date %>% 
  dplyr::count(site)

lobster_site
```

If we want to create a summary table that contains statistics OTHER than counts by group, it's easier to use group_by() + n()

```{r}
lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    lobster_number = n()
  )

lobster_summary
```

#### Find confidence intervals

Use the t.test() function to find confidence intervals (for one sample) and perform t-tests to compare means of two samples (...this will be covered conceptually in lectures week 8)
```{r}
ivee_lobster <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

# What is vector of observations? And a t-test I really should eat before doing this lab. 

t.test(ivee_lobster)

#null hypothesis= most boring hypothesis between means
# for example, if you have the most perfectly randomly drawn samples, their differences should be 0.
# Since it is not 0, then it is significantly different
```

#### Two-sample t-test to compare means

We want to ask: Is there a significant difference in lobster lengths at Naples and Mohawk reefs? 

We've done our necessary exploratory analyses to determine that a 2-sample t-test for means comparison is appropriate.

This is a ranked based test. 

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mn_ttest <- t.test(mohk_sample, napl_sample)

mn_ttest
```

There is a significant different in lobster lengths between Naples and Mohawk reef (t(df) = statistic, p < .001, alpha = .05). 

```{r}
lobster_mn <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

mn_ttest2 <- t.test(size_mm ~ site, data = lobster_mn)

mn_ttest2

mn_ttest2$p.value
mn_ttest2$statistic
```

There is a sig diff (t(1850.8) = -19.849, p = 2.2e-16). 

The above is not a good line because it is not reproducible if you have to copy and paste all of the results and if you have to change something in the data before this line of text, then your results won't work anymore. 

Do this: 

Here is the p-value: `r mn_ttest2$p.value`

Here is my t-statistic: `r mn_ttest2$statistic`

#### Now: a heatmap

```{r}

lobster_ys <- lobster_parse_date %>% 
  count(obs_year, site)

ggplot(data = lobster_ys, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

